---
title: "Learning2Code"
author: "Steph"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Basics

Slide 36:

```{r cars, echo=TRUE}

He_amu <- 4.0026

#This is the atomic mass of Helium in amu


print(He_amu)

He_amu
```

## Data types

Slide 37:

```{r, echo=TRUE}
He_amu <- 4.0026
#numeric
He_amu

# for writing notes

He_amu <- 4L
#integer
He_amu

He_amu <-"4.0026"
#character
He_amu

He_amu_is_4.0026 <- TRUE
#logical data 
He_amu_is_4.0026
```

#### why is data type important?

Slide 38:

```{r, echo=TRUE}

as.integer(30/4)
```

### Data structures

Slide 40:

```{r pressure, echo=TRUE}
elements <- c("H","He","Li","Be","B")
elements

list <- list("hello" , 1, TRUE, 4L, FALSE)
list

String <-"Periodic Table"
String <- paste("Periodic", "Table",sep= " ")

amu <- c(1.008, 4.003, 6.941, 9.012, 10.811) 

dataframe <- data.frame(elements,amu)
dataframe


library(tidyverse)

tibble <- as_tibble(dataframe)
tibble


```

### What is the atomic mass of Beryllium?

Slide 42:

```{r pressure, echo=TRUE}

tibble[4,]
#outputs the 4th row

tibble$amu[4]
#outputs the 4th row of column "amu"

tibble[which(tibble$elements == "Be"),]
#searches for which row in column "elements" meets the condition "Be"

elements[4]

```

### What is the molecular mass of LiH and H2?

Slide 43:

```{r math}



```

### Conditionals

#### Navigating binary decision trees

Slide 44:

```{r}

```

### Loops

#### How do we differentiate for and while loops?

Slide 46:

```{r}
for(i in 1:3){
  print(tibble[i,])
}
i=1
while (i <= 3){
  print(tibble[i,])
  i= i + 1
}
```

### Functions

#### Create a function to convert temperature units

Slide 48:

```{r}
Cel2Kel <- function(temp){
  kelvin = temp + 273.15
  return(kelvin)
}
Cel2Kel(1)

Kel2Cel <- function(temp){
  Celsius = temp - 273.15
  return(Celsius)
}
Kel2Cel(274.15)
```

# Packages: Tidyr and Dplyr

Begins on slide 51 on the slide deck

### How-to import data

Slide 53:

```{r}
amu <- read.csv("Data - week 1 and 2.csv")
```

### Calculate the mass number from our imported data

Slide 56

```{r}
library(tidyverse)

amu <- amu %>%
  mutate(massnumber = atomic.number + neutrons)

```

### Calculate the average atomic mass from our imported data

Slide 57

```{r}
amu <- amu %>%
  group_by(symbol) %>%
  summarise(av.mass = sum(abundance/100*mass))
```

### Calculate the molar mass of iron (II) chloride

Slide 60-62:

```{r}
#potential sol 1(brute force)
FeCl2.mass <- as.numeric(amu[3,2]*2 + amu[4,2])
#potential sol 2
FeCl2.mass <- amu %>%
  filter(symbol == "Fe" | symbol == "Cl") %>%
  arrange(symbol)

FeCl2 <- as.numeric(FeCl2.mass[2,'av.mass'] + FeCl2.mass[1, 'av.mass']*2)

#SOl 3
FeCl2.mass <- as.numeric[which(amu$symbol == "Fe"), 'av.mass'] + amu[which(amu$symbol == "Cl"), 'av.mass']*2
```

## Challenge: can we write a function that can calculate the molar mass of any molecule written as a molecular formula

Potential solution for challenge (slide 63)

```{r}
MMass_lookup <- setNames(amu$symbol, amu$av.mass)

MMass_Calc <- function(molecule){
  matches <- gregexpr("[A-Z][a-z]?[0-9]*", molecule)
  #^looks for a capital letter then maybe a lowercase letter (?) to parse elements in the formula, then a number behind the letter to determine how many of each element there is
  parts <- regmatches(molecule, matches)[[1]]
  #^finds individual elements identified in previous function and defines them in new dataset parts
  total <- 0

  for (part in parts) {
    element <- sub("[0-9].*","",part)
    count <- sub("[A-Za-z]+","",part)
    
    if (count == "")  count <- 1 
    
     else count <- as.numeric(count)
    
    total <- total + amu[which(amu$symbol==element),'av.mass'] * count
  }
  
  return(total)
  }
  
MMass_Calc("C2H6")
  
  


```

# Making graphs in R

Starting from Slide 65

```{r}

```

Slide 70

```{r}

  
```

## Challenge: can we identify the individual elements on the plot?

Adding more information to the graphs, examples (slide 73):

```{r}

  
```

## Least squares linear regressions

Slide 75

```{r}


```

You can also extract the information without plotting it - Slide 76

```{r}

```
